---
title: "Data Tidying"
author: "Eloise Newman"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, true = FALSE)
library(tidyverse)
```

Data tidying of raw data for "Hot and Bothered" masters research. R version is `r R.Version()$version.string`. Data wrangling involves mainly using the `dplyr` package from `tidyverse` by Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2022).

# Three raw data sets

Three data sets imported to tibbles:

1.  **flymass** - dry fly mass in micrograms (mg) using the Mettler Toldero microbalance
2.  **meta_data** - contains treatment for each file number, time, emergence, starvation date etc.
3.  **behaviours** - cursed output from BORIS via JWatcher

```{r include=TRUE, warning = FALSE, message=FALSE}
# fly mass dataset
flymass <- read_csv("data/flymass.csv")

# fly video dataset (temp, time, eclosion time)
meta_data <- read_csv("data/Fly Video List.csv", n_max = 38, col_select = ...1:PANstarvTime)

# fly behaviours dataset (from Boris or Jwatcher)
behaviours <- read_csv("data/summary.csv", skip = 9)
```

# Data tidying and wrangling

## Fly mass

None

## Fly video meta-data (treatment, time etc)

```{r Tidying up the Fly video meta data set, include=TRUE}
# rename columns
meta_data2 <- meta_data %>% 
  dplyr::select(-Arena_Start, -Arena_End) %>% 
  rename(date_experiment = `...1`, )

## Need to make Temp a ordered factor with levels
temp_levels <- c("L", "M", "H")
meta_data2$Temp <- factor(meta_data2$Temp, levels =  temp_levels)
```

Further potential changes not done:

1.  convert egg collection (ec) and emergence (em) date to days
2.  calculate days from emergence to fight, emergence to starv, s
3.  recalculate file_end_time, and change column to time

## Fly behaviour data set from BORIS/Jwatcher

Renaming the columns into something recognisable

```{r Renaming columns for fly behaviour, include=TRUE}
beh_columns <- colnames(behaviours)

  # renaming Prefixes
beh_columns <- str_replace(beh_columns, "StateAllInt ", "Int") 
beh_columns <- str_replace_all(beh_columns, "StateAllDur ", "Dur")

  # Rewnaming Behavioural codes for behaviours
beh_columns <- beh_columns %>% 
      # Prefixes
          str_replace("cd.res file", "Fly_Number")  %>%
          str_replace_all("StateAllInt ", "Int")  %>%
          str_replace_all("StateAllDur ", "Dur") %>%
      # PAL
          str_replace_all("-0-\\!", "WingThreatPAL") %>%
          str_replace_all("-0-\\@", "LungePAL")  %>%
          str_replace_all("-0-A", "MovingPAL")   %>%
          str_replace_all("-0-C", "LegShovePAL")   %>%
          str_replace_all("-0-D", "BodyShovePAL")   %>%
          str_replace_all("-0-E", "HeadbuttPAL")   %>% 
          str_replace_all("-0-S", "StopMovePAL")   %>%
          str_replace_all("-0-R", "RetreatPAL")   %>%
          str_replace_all("-0-Q", "FeedPAL")   %>%
          str_replace_all("-0-W", "StopFeedPAL")   %>%
          str_replace_all("-0-X", "StopFencePAL")   %>%
          str_replace_all("-0-Z", "FencingPAL") %>% 
      # PAN   
          str_replace_all("-0-1", "WingThreatPAN")  %>%
          str_replace_all("-0-2", "LungePAN")  %>%
          str_replace_all("-0-a", "MovingPAN")   %>%
          str_replace_all("-0-c", "LegShovePAN")   %>%
          str_replace_all("-0-d", "BodyShovePAN")   %>%
          str_replace_all("-0-e", "HeadbuttPAN")   %>%
          str_replace_all("-0-s", "StopMovePAN")   %>%
          str_replace_all("-0-r", "RetreatPAN")   %>%
          str_replace_all("-0-q", "FeedPAN")   %>%
          str_replace_all("-0-w", "StopFeedPAN")   %>%
          str_replace_all("-0-x", "StopFencePAN")   %>%
          str_replace_all("-0-z", "FencingPAN")

# Reassign to behaviours data set
colnames(behaviours) <- beh_columns
```

Finally need to remove the suffix from fly number

```{r Removing suffix from fly number in fly behaviour, include=TRUE}
#Remove "_No focal subject.cd.res" from fly number
behaviours <- behaviours %>%
  separate(Fly_Number, c("Fly_Number", "redunt"), sep ="_No " ) %>% 
  dplyr::select(-redunt)
```

### Behaviours and codes - Prefixes

[***ref***]{.underline}***: CHAPTER 10 Basic Analysis: Interpreting Your Output - JWatcher Manual***

-   ***StateAllInt-\> Int** refers to "State Analysis, All, Interval" statistics. The statistics are calculated for the intervals between states, regardless of whether they begin and end completely with- in a time bin.*

-   ***StateAllDur** -\> **Dur** refers to "State Analysis, All, Duration" statistics. The statistics are calculated for the durations of states. All durations of a behavioral state are included in the calculations, regardless of whether they begin and end completely within a time bin, and regardless of whether they occur adjacent to the onset or termination of the out- of-sight key. In this case, the out-of-sight key affects the proportion of time-in-sight statistic only.\**

| Code | Meaning                           |
|------|-----------------------------------|
| N    | Occurrence                        |
| TT   | Total Time (milliseconds)         |
| X    | Average (milliseconds)            |
| SD   | Standard deviation (milliseconds) |
| Prop | Proportion of time                |

### Behaviours selected

StateAllDur N-0 \# the number of bouts of that behaviours (both PAN & PAL)

-   wing threat

-   lunge

-   leg shove

-   body shove

-   headbutt

-   retreat

-   feed

-   moving

-   fencing

StateAllDur TT-0 \# the total duration of that behaviour recorded in that observation.

-   TT fencing duration

-   TT moving duration

-   TT feeding

StateAllDur X-0

-   average fencing duration

-   average moving duration

-   average feeding

### Selecting columns

and converting time from milliseconds to seconds

```{r Selecting the columns I am interested in, include=TRUE}
# Selecting the columns I am interested in
selectbehaviour <- behaviours %>%
  dplyr::select(Fly_Number, 
         # number of bouts
         `DurNBodyShovePAL`, `DurNBodyShovePAN`,
         `DurNFeedPAL`, `DurNFeedPAN`,
         `DurNHeadbuttPAL`, `DurNHeadbuttPAN`,
         `DurNLegShovePAL`, `DurNLegShovePAN`,
         `DurNLungePAL`, `DurNLungePAN`,
         `DurNWingThreatPAL`, `DurNWingThreatPAN`,
         `DurNRetreatPAL`, `DurNRetreatPAN`,
         `DurNMovingPAL`, `DurNMovingPAN`,
         `DurNFencingPAL`, `DurNFencingPAN`,
         # total duration
         `DurTTFeedPAL`, `DurTTFeedPAN`,
         `DurTTMovingPAL`, `DurTTMovingPAN`,
         `DurTTFencingPAL`, `DurTTFencingPAN`,
         # average duration
         `DurXFeedPAL`, `DurXFeedPAN`,
         `DurXMovingPAL`, `DurXMovingPAN`,
         `DurXFencingPAL`, `DurXFencingPAN`,
          ) %>% 
  mutate( #PAL total
      DurTTFencingPAL = DurTTFencingPAL/1000,
      DurTTMovingPAL = DurTTMovingPAL/1000,
      DurTTFeedPAL = DurTTFeedPAL/1000,
         # PAN TOTAL
      DurTTFencingPAN = DurTTFencingPAN/1000,
      DurTTMovingPAN = DurTTMovingPAN/1000,
      DurTTFeedPAN = DurTTFeedPAN/1000,
          # PAL average
      DurXFencingPAL = DurXFencingPAL/1000,
      DurXMovingPAL = DurXMovingPAL/1000,
      DurXFeedPAL = DurXFeedPAL/1000,
         # PAN Average
      DurXFencingPAN = DurXFencingPAN/1000,
      DurXMovingPAN = DurXMovingPAN/1000,
      DurXFeedPAN = DurXFeedPAN/1000)
```

Using `readr` package from `tidyverse` to export as a tidied data file.

```{r exporting tidied fly behaviour data to csv, include=TRUE}
write_excel_csv(selectbehaviour, "tidy_data/behavioural_data_tidied.csv", col_names = TRUE)
```

# Combining data sets

## Behaviours and mass - `behaviour_mass`

First I combine the fly mass to fly behaviour dataset by unique `Fly_Number` to a dataset called `behaviour_mass`

```{r Fly fight data and fly mass data}
# USING THE FLY NUMBER or FILE NAME, Assign fly mass to fly behaviour dataset 
behaviour_mass <- 
  left_join(selectbehaviour, flymass, by = "Fly_Number")
```

## Behaviour, mass and meta data - `behaviour_mass_meta`

Next I assign the meta data (aka the treatment and emergence time etc) to the fly behaviour + fly mass data set (`behaviour_mass_meta`)

```{r Fly meta data and mass and fight data}
# Assigning meta data to fly fights + fly mass. Meta_data adds an NA row so removing 
# that as well
behaviour_mass_meta <- behaviour_mass %>% 
  mutate(file_name = Fly_Number) %>% #generating file_name from fly number
  separate(file_name, c("file_name", "remove"), sep = 3) %>% # to match to meta
  dplyr::select(-remove) %>% 
  right_join(meta_data2, by = "file_name") %>%
  filter(!is.na(Fly_Number))
```

Therefore, in the `behaviour_mass_meta` data set, each row is one fight with behaviour, mass and metadata for PAN and PAL in one row - species are combined.

## Meta data and mass - `flymass_meta`

Separately I join some of the `meta_data` to the `flymass` dataset to be able to separately analyse fly mass to temperature. I also convert to long format so that each fly has their own row, and a species column is added.

```{r Fly video data + fly mass data}
# Adding fly mass + metadata togther in wide format
flymass_meta <- flymass %>% 
  mutate(file_name = Fly_Number) %>% #generating file_name from fly number
  separate(file_name, c("file_name", "remove"), sep = 3) %>% #to match to meta
  dplyr::select(-remove) %>% 
  right_join(meta_data2, by = "file_name") %>% 
  mutate(mass_dif = PAL_Mass_mg - PAN_Mass_mg) # mass difference between opponents

    ## need to convert to long format from wide
flymass_meta <- flymass_meta %>% 
  pivot_longer(c(PAN_Mass_mg, PAL_Mass_mg), names_to = "Species", values_to = "Mass") %>% 
  separate(Species, c("Species", "remove"), sep = 3) %>% 
  dplyr::select(Fly_Number, file_name, Letter, Sex, Temp, temp_room, 
         Species, Mass, mass_dif, -remove) %>% 
  filter(!is.na(Mass)) %>% #converting zeros to NA to make easier to filter
  na_if(0.000)
```

## Separating PAN and PAL behavioural data into separate rows - `pan_pal`

Effectively making this in long format.

```{r}
# Separate PAN and PAL data + add a column intitled `Species`
PAN <- behaviour_mass_meta %>% 
  dplyr::select(Fly_Number, contains("PAN"), 
         file_name, Temp, Sex, ECDate, EmDate, temp_room, file_start_time) %>%
  mutate(Species = "PAN") 

PAL <- behaviour_mass_meta %>% 
  dplyr::select(Fly_Number, contains("PAL"), 
         file_name, Temp, Sex, ECDate, EmDate, temp_room, file_start_time) %>%
  mutate(Species = "PAL") 

# combine them together - ignoring the PAL column names
      PAL_pan_names <- PAL
      names(PAL_pan_names) <- names(PAN)
pan_pal <- rbind(PAL_pan_names, PAN)

# Rename/remove the "PAN"
col_names <- names(pan_pal)
  col_names <- col_names %>% 
    str_replace("PAN", "")
names(pan_pal) <- col_names # reassign

# Select
pan_pal <- pan_pal %>% 
  dplyr::select(1:16, Species) 

# Re-align + connect metadata and mass as janky
pan_pal <- 
  left_join(pan_pal, flymass_meta, by = c("Fly_Number", "Species")) %>% 
  mutate(pan_pal, fencing_presence = ifelse(DurTTFencing > 0, 1, 0))
```

Also get one dataset for just PAN and PAL respectively.

```{r}
# reassign species specific datasets to species
PAN <- pan_pal %>% 
  filter(Species == "PAN")
PAL <- pan_pal %>% 
  filter(Species == "PAL")
```

# Original analysis datasets

## Summarising total acts of aggression - `pan_pal`

```{r total aggressive acts by counting incidence + counts, message=FALSE, warning=FALSE}
pan_pal <- pan_pal %>% 
  rowwise() %>% #total ag acts for each row otherwise shenanigans
  mutate(total_ag_acts = sum(`DurNBodyShove`, #total ag acts for each row
                               `DurNHeadbutt`,
                               `DurNLegShove`,
                               `DurNLunge`,
                               `DurNWingThreat`,
                               `DurNFencing`)
       )

# Pivoting pan_pal dataset for each act = pan_pal_ag
pan_pal_ag <- pan_pal %>% 
  pivot_longer(c(2, 4:7, 10), names_to = "ag_act", values_to = "Count") %>% 
 dplyr::select(Fly_Number, file_name, Letter, Sex, Temp, Species, 
        DurTTMoving, DurTTFeed, DurTTFencing,
         Mass, mass_dif, total_ag_acts, ag_act, Count) 
```

## Calculating winners and losers of a bout - `winners_losers_pair`

Do not have a "win" count in analysis, only retreat and number of fencing bouts. Also have duration of fencing (how hard fought). Therefore retreats and number of bouts was used to calculate wins. Number of retreats is the number of loses.

```{r working out the losers intial}
winners_losers_pair <- behaviour_mass_meta %>% 
  mutate(losesPAL = DurNRetreatPAL, # Proportion of number of personal loses
         losesPAN = DurNRetreatPAN,
         drawsPAL = DurNFencingPAL-DurNRetreatPAL-DurNRetreatPAN,
         drawsPAN = DurNFencingPAN-DurNRetreatPAL-DurNRetreatPAN,  
         winsPAL = DurNFencingPAL-drawsPAL-DurNRetreatPAL,
         winsPAN = DurNFencingPAN-drawsPAN-DurNRetreatPAN,
         mass_dif = PAL_Mass_mg-PAN_Mass_mg,
         bouts = DurNFencingPAL,
         resolved = DurNFencingPAL-drawsPAL) %>% 
  dplyr::select(Fly_Number, Letter, Temp, Sex, temp_room, bouts,
         PAN_Mass_mg, PAL_Mass_mg, mass_dif, 
         DurTTFencingPAN, DurTTFencingPAL, resolved,
         losesPAL, losesPAN, drawsPAL, drawsPAN, winsPAL, winsPAN)
```

Data wrangling so PAN and PAL are in each rows... could just start with pan_pal but oh well!

```{r matching meta data with winenrs losers}
# Separating out Pan and Pal mass
winners_losers <- winners_losers_pair %>% 
  separate(PAN_Mass_mg, c("PAN", "removePAN"), sep = 6) %>% #to match to meta
  separate(PAL_Mass_mg, c("PAL", "removePAL"), sep = 6) %>% #to match to meta
  dplyr::select(-removePAN, -removePAL) %>% 
  pivot_longer(c(PAN, PAL), names_to ="Species", values_to = "Mass")

### separate by species and calulate total bout duration
winners_losers_PAN <- winners_losers %>%  
  dplyr::filter(Species == "PAN") %>% 
  dplyr::select(-losesPAL, -drawsPAL, - winsPAL, -DurTTFencingPAL) %>% 
  dplyr::rename(loses = losesPAN, draws = drawsPAN, wins = winsPAN,
                TTBouts = DurTTFencingPAN)
              # rename columns with no PAN and PAL
winners_losers_PAL <- winners_losers %>%  
  dplyr::filter(Species == "PAL") %>% 
  dplyr::select(-losesPAN, -drawsPAN, - winsPAN, -DurTTFencingPAN) %>% 
  dplyr::rename(loses = losesPAL, draws = drawsPAL, wins = winsPAL,
                TTBouts = DurTTFencingPAL)

## join the two together, no by as same column names
winners_losers <- 
  full_join(winners_losers_PAN, winners_losers_PAL) 

# bouts dataset with zero bouts
winners_losers_bout <- winners_losers %>% 
pivot_longer(c(loses, draws, wins), # all together
                names_to = "outcome", values_to = "result") %>% 
  filter(!is.na(result)) # remove NA

# bouts dataset with no bouts removed
winners_losers_bout_removed <- winners_losers_bout %>% 
  filter(bouts > 0) # remove NA
```

# List of generated datasets

This code should have generated this in the enviroment, not all needed for analysis.

Original imported, untidied raw data:

1.  `behaviours` - behavioural data from BORIS/JWatcher

2.  `flymass` - dry fly mass in micrograms (mg)

3.  `meta_data` -contains treatment for each file number, time, emergence, starvation date etc.

And this code has generated:

1.  `behaviour_mass_meta` - each row is one fight with behaviour, mass and metadata for PAN and PAL in one row.

2.  `flymass_meta` - mass and metadata where each row is a fly.

3.  `pan_pal` - each row for each fly, including total ag_acts and presence of fencing.

    1.  `pan_pal_ag` - each row for each type of ag_act, so 6 rows for each fly.

4.  `selectbehaviour` - tidied full dataset with selected behaviours.

5.  `meta_data2` - tidied metadata

6.  `winners_losers` - has win, loss, draw for each fly separately

    1.  `winners_losers_pair` - has wins,losses, draws for each fight and species one row. aka paired

    2.  `winners_losers_PAN` or `_PAL` - separated by species

    3.  `winners_losers_bout` - pivots, so outcome is draw, loss, win as one column with result. Each fly has three rows.

    4.  `winners_losers_bout_removed` - dataset removing fights with no bouts recorded.

For the analysis, some of these datasets are required. Either run the script above to have them in your environment. The code below places copies in the data folder.

```{r exporting csv files for analysis}
write_csv(pan_pal, "tidy_data/pan_pal.csv", col_names = TRUE)
write_csv(flymass_meta, "tidy_data/flymass_meta.csv", col_names = TRUE)
write_csv(winners_losers, "tidy_data/winners_losers.csv", col_names = TRUE)
write_csv(winners_losers_pair, "tidy_data/winners_losers_pair.csv", col_names = TRUE)
```

# Colours and theme

## Colour scheme

```{r masters colour scheme hex codes}
masters_colours <- c("#264653", #Charcoal - colour used for PAL
                     "#2a9d8f", #Persian Green  - win or low temp
                     "#e9c46a", #Maize Crayola
                     "#f4a261", #Sandy Brown - draw or medium temp
                     "#ef626c", #Fiery Rose - loss or high temp
                     "#B75688" # Mulberry - colour used for PAN
                     )
```

## Theme for graphs

Theme setting for graphs used in ggplot2 - just use + and theme_masters.

```{r Creating theme for masters}
theme_masters <- theme_bw() + theme( 
   plot.title = element_text(hjust = 0.5, size = 18),
   text = element_text(size = 12),
   axis.text = element_text(size = 12),
          axis.title = element_text(size = 16),
   legend.text = element_text(size = 12),
        legend.background = element_rect(fill="white"),
        legend.position = "bottom",
    panel.grid = element_line(color = "gray94"), #grid
    strip.background = element_blank(),
          strip.text = element_text(size=11, face="bold")
   )

theme_set(theme_masters)
```
